### 2019-05-26 paper/paper.tex
1

\documentclass[runningheads]{llncs}
\usepackage{graphicx}
\usepackage[width=122mm,left=12mm,paperwidth=146mm,height=193mm,top=12mm,paperheight=217mm]{geometry}
\usepackage{amsmath,amssymb} % define this before the line numbering.

\DeclareMathOperator*{\argmin}{argmin}

%\usepackage{ruler}
\usepackage{color}

%\documentclass{llncs}
%\usepackage{graphicx}
%\usepackage{geometry}
%\geometry{margin=2cm}
%\usepackage{amsmath}

\usepackage{pgfplots}
\usepackage{filecontents}
\usepackage[caption=false]{subfig}
\usepackage{listings}
\lstset{
basicstyle=\small\ttfamily,
columns=flexible,
breaklines=true
}

\begin{document}
\title{Dissecting Neural Style Transfer}
\author{Xi Du}
\institute{Australian National University, Australia\\
\email{u6559090@anu.edu.au}}
\maketitle 
\begin{abstract}
\cite{method}

\keywords{Neural network, Deep learning, 
Small data, Interpretability .}
\end{abstract}

\section{Introduction}

\section{Literature Review}

\section{Method}

\subsection{Deep Neural Network}

For the purpose of implementing and even extending neural style transfer algorithms,
it is not necessary to understand how a neural network is trained, because
we can use pretrained model for example vgg19 \cite{vgg19}.

It is necessary to point out that a deep neural network that neural style transfer
is concerend about
is just a sequence of functions
$F_1,F_2,F_3,...$
whose inputs and outputs are all multi-dimensional arrays, as illustrated in Figure \ref{f1f2f3}.
\begin{figure}
\center
\includegraphics[width=0.6\textwidth]{f1f2f3.pdf}
\caption{A deep neural network \label{f1f2f3}}
\end{figure}
Because the arrays have usually more than two numbers of dimensions
they are not really matrices.
The arrays are called tensors (as in TensorFlow or torch.Tensor in PyTorch) 
but actual tensor algebra in a mathematical sense is rarely relevant.
The functions need to be differentiable, which is handled automatically by modern 
deep learning frameworks.
Other than that we can treat the functions as blackboxes because we are 
not concerned of training the model.
A ``layer'' in a deep neural network refers to a few consecutive functions together.
It is not particularly necessary to consider ``layers'' in this work. 
Though it is important to note that in our illustrations there are functions instead of layers.

\subsection{vgg19 model}
For the vgg19 model that most neural style transfer implementations available are based on,
The inputs and outpus of each function are all arrays with 4 dimensions.
The word ``dimension'' here may mean something slightly different from what ``dimension'' means in
for example ``$3$-dimension vector''. 
Some people call the number of dimensions ``rank'' which would then raise another confusion with the rank of matrices.
We give an illustration for the vgg19 model in Figure \ref{vgg512}
\begin{figure}
\center
\includegraphics[width=\textwidth]{vgg512.pdf}
\caption{The VGG19 model with an $512\times512$ RGB image as its input \label{vgg512}}
\end{figure}

The 1st dimension is the ``batch'' dimension, which means that you can put several images through the
series of functions.
In neural style transfer implementations it is usually just one image each time.
So the 1st dimension is always 1, even between the $F$ functions, throughout this work.

The 2nd dimension is the ``feature'' dimension. For a raw RGB image, it is 3.
Intermediate arrays between, for example, $F_3$ and $F_4$, usually 
have a size much larger than 3, such as 64 or 128.

The 3nd and 4th dimensions are just spatial locations. 
For a $512\times512$ image they are 512 and 512. 
For intermediate arrays, these 2 dimensions are sometimes scaled down
to half or even $\frac{1}{4}$ of the sizes of the image.

The functions $F_1,F_2,F_3,...$ now have meaningful names such as relu\_1, conv\_2 now.
[[[[Not actually necessary to know but insert your elaborations here]]]]

The arrays after the initial image are called feature maps, because their value means
``how much a feature x exists at position y and x''.
For example in a $1\times64\times512\times512$ feature map, it means there are 64 types of features.
and its [1, 17, 111, 222] element, would then refer to the extent that the 17th type feature exists
at the position of the 222th column of the 111th row. This ``extent'' could also be negative.

\subsection{Vanilla neural style transfer}

The vanillay neural style transfer algorithm \cite{nst} is 
structured as iteratively solving an optimisation problem.
The argument to optimise is the image as a multidimensional array,
of size $1\times3\times512\times512$ for example.
The interative solver can be many, but in our case it is 
L-BFGS [[[[Maybe elaborate on L-BFGS here]]]]. 
The initial value of the argument could be either white noise or the content image,
but the latter appeared to make the optimisation much easier.

The key issue here is still how to structure the optimisation target.
The value to minimise is a linear combination of a ``style loss'' 
and ``content loss''.
Although these appeared to be two weights,
actually only their ratio mattered.
We simply fixed the weight of the content loss to 1 and
leave the weight of the style loss as a adjustable hyperparameter $k$.
\begin{align}
\argmin\limits_{\text{image}} (kL_\text{style}+L_\text{content})
\end{align}

The content loss $L_\text{content}$ is the easier one to explain of the two losses.
It is the mean square error between the outputs of a certain function in the
vgg19 taking the current argument image and the content image as inputs respectively.
We say ``a certain'' function because the output of which function to choose is
adjustable.
A typical choice is the output of \verb|conv_4| function. See Figure \ref{contentloss}.
\begin{figure}
\center
\includegraphics[width=\textwidth]{contentloss.pdf}
\caption{The content loss \label{contentloss}}
\end{figure}

The style loss $L_\text{content}$ can actually be a sum of several style losses, or sometimes just one.
For each style loss, 
both the current image under optimisation and the style image are passed through the vgg19 model.
The two outputs of a certain function in the model are extracted, 
and converted to Gram matrices. We will elabrate on what Gram matrices are,
for now they are just square matrices computed from multidimensional arrays (i.e. feature maps).
For example, the Gram matrix (in the style transfer algorithm) of a $1\times64\times512\times512$ array
is a $64\times64$ matrix.
The point is to compute something that describe the distributions of features
in the image without the spatial information of where the features are.
Gram matrices describe the distributions with correlation between features.
The several style losses are computed with the outputs of different functions.
Otherwise, each of the several style losses are computed in exactly the same way.
Again outputs of which functions in the model to derive style losses from is really adjustable.
A nice default is \verb|conv_1|,\verb|conv_2|,...,\verb|conv_5|. We illustrate
the style losses derived from the output of \verb|conv_1| in Figure $\ref{styleloss}$.

\begin{figure}
\center
\includegraphics[width=\textwidth]{styleloss.pdf}
\caption{The style loss derived from the output of conv\_1 \label{styleloss}}
\end{figure}

\subsection{Gram matrix}
[[[[Elaborate on gram matrix here]]]]

\subsection{Removing the spatial information in other ways}
We think that the most brilliant aspect of the neural style transfer algorithm,
apart from the general structure, is the use of Gram matrices to compare 
distributions of features in two images, so that where the features are does not matter.
It gives really nice insight on what ``artistic style'' means.

Now we wonder, could we achieve similar results with other ways to compare 
distribution of features?
One approach would be to consider the \emph{comparison} as a whole to find 
substitutes. 
For example, minimising the mean square errors between two Gram matrices could be seen as 
a special case for minimising a Maximum Mean Discrepancy, which means other Maximum Mean Discrepacies
could be used \cite{MMD}.
For us, it seemed more natural to consider substitutes for the process itself of 
transforming a feature map into a Gram matrix. 
That is, we will consider other
mappings that remove the spatial information of where features are but retains
the distribution of features, like a histogram.
We will call such functions ``\emph{dislocators}'' in the remaing part of this work.
Unfortunately, the word ``dislocate'' would give the connotation of moving something to somewhere else.
However, the word ``locator'' means something that provides spatial information,
so calling somthing that removes spatial information a ``dislocator'' is still plausible.

Some ``dislocator'' functions we tried are

\section{Experiments and Discussion}

\subsection{Varying the dislocator function}

\subsection{Varying the extraction points for losses}

We will also show that other factors, such that the outputs of which functions to derive the style 
losses from are not very essential and somewhat belongs to tuning.

\section{Conclusion and Future Work}

\begin{thebibliography}{8}
\bibitem{nst}
Gatys, L., Ecker, A., \& Bethge, M. (2016). A neural algorithm of artistic style. Journal of Vision, 16(12), 326. doi:10.1167/16.12.326
\bibitem{MMD}
Li, Y., Wang, N., Liu, J., \& Hou, X. (2017). Demystifying Neural Style Transfer. IJCAI.
\bibitem{interp}
Doshi-Velez, F., \& Kim, B. (2017). Towards A Rigorous Science of Interpretable Machine Learning.
\bibitem{pytorch}
Paszke, A., Gross, S., Chintala, S., Chanan, G., Yang, E., DeVito, Z., Lin, Z., Desmaison, A., Antiga, L., \& Lerer, A. (2017). Automatic differentiation in PyTorch.
\end{thebibliography}
\end{document}