### 2019-06-04 paper/paper.tex
Updates on introduction and related work by Jianzhen

\documentclass[runningheads]{llncs}
\usepackage{graphicx}
\usepackage[width=122mm,left=12mm,paperwidth=146mm,height=193mm,top=12mm,paperheight=217mm]{geometry}
\usepackage{amsmath,amssymb} % define this before the line numbering.

\DeclareMathOperator*{\argmin}{argmin}

%\usepackage{ruler}
\usepackage{color}

%\documentclass{llncs}
%\usepackage{graphicx}
%\usepackage{geometry}
%\geometry{margin=2cm}
%\usepackage{amsmath}

\usepackage{pgfplots}
\usepackage{filecontents}
\usepackage[caption=false]{subfig}
\usepackage{listings}
\lstset{
basicstyle=\small\ttfamily,
columns=flexible,
breaklines=true
}

\begin{document}
\title{Dissecting Neural Style Transfer}
\author{Xi Du}
\institute{Australian National University, Australia\\
\email{u6559090@anu.edu.au}}
\maketitle 
\begin{abstract}
\cite{method}

\keywords{Neural network, Deep learning, 
Small data, Interpretability .}
\end{abstract}

\section{Introduction}

Image stylization is the process that mimics the appeal of artistic media [8] 
and converts input images into a stylized version. It is a research started with
semi-automatic painting systems since 1990s [1] and can be used in many newly 
developed applications in machine learning and artificial intelligence [5]. 
This technique leads to a new form of communication [2]. With the use of image 
stylization, the media users can keep stylistic consistency on media platforms 
like Instagram or Twitter without obtaining skills in traditional artistic 
techniques [16]. Image stylization allows users to generate picture in 
traditional artistic styles, which had been expensive and time-consuming with 
traditional techniques [19]. It shifts the way how the media users share 
pictures on internet. At the beginning of the image stylization, which it 
started in 1980s, procedural filters and patch-based algorithms were studied 
and applied [2]. One example for the procedural filters is to refine image 
with strokes placement algorithms, which a grid of rendering brush strokes is 
placed over an image [13, 3]. Later on, patch-based synthesis was studied based 
on texture recognition [4]. Moreover, these two methods have numerous 
difficulties in further developments on the capture of complex nuances of an 
artist’s style, and can only work for certain types of images [2]. However, 
deep neural networks, which are developed as a general-purpose tool working 
different computer vision problems [5].  The paper published by Leon Gatys in 
2015 has proven that a simple neural network works well on the texture 
recognition just like the patch-based synthesis algorithms [14]. 
One of the applications is auto-painter for cartoon image generation and it 
applies deep neural networks on image stylization [11]. These computational 
bio-inspired neural networks mimic the basic structure of brain neurons. 
Like the auto-painter, convolutional neural network is used in this paper 
for developing a tool in image stylization. Neural networks, especially for 
the deep artificial neural networks, has been widely studied and used in the 
pattern recognition and machine learning in recent years [17]. Convolutional 
neural network is one of the deep learning neural networks and have 
substantially achieved the state-of-the-art accuracies of object recognition 
[12]. This is the key in image stylization where the essential features in the 
style image are extracted and applied on the input image.

This paper applies a pre-trained CNN, VGG-19, on image stylization and 
extends its possibilities on image stylization. CNNs have a similar structure 
as ordinary neural networks, which include neurons with learnable weights and 
biases. CNNs are typically feed-forward architecture and widely used in computer
vision tasks in recent years [12]. Its basic unit is single neuron. Each of 
these neurons takes one or multiple inputs, conducts a dot product then 
optimally follows it with a non-linearity [9]. The core mathematical equation 
for CNN is a result derived from two functions by integration and this indicates 
how one’s shape is changed by the other. CNN works in a way similar to broad 
categorization [19]. When a CNN is used in image processing, every pixel of the 
input image will be processed by a feature detector and a feature map is given 
as the result. There will be very less, but critical features kept in the 
feature map compared to the input image. Consequently, some key features might 
be missed, and massive trainings are required for CNN’s accuracy on texture 
capture.

\section{Related Work}

In the field of image stylization, there are abundant related work found.

2.1 Image stylization using stroke-based and example-based methods

Both stroke-based and example-based methods have been studied since last century 
for image stylization. For the stroke-based stylization, one typical method is 
that brush strokes are aligned based on input image for varying color, size and 
orientation [18]. Many methods, which are developed based on brush strokes, 
normally either focus on one specific type of brush or provides a generic 
framework [7]. For an instance, a brush stroke for a traditional Japanese 
sumi-e style painting creation is modelled with the use of a 1D array of 
bristles and this method can only generate a stylized image in grey scale [20]. 
One major disadvantage of the stroke-based method is that it cannot generate 
result automatically and provides tools to users as they can tweak the output 
[7]. Many researches have also pointed out that the complex nuances of an 
artist’s style are too hard to be captured in some degrees [2]. As for the 
example-based methods, texture recognition and transfer are the core parts [18].
One framework for this method is called the image analogies, which is a basic 
reasoning process and consisted of two stages: a design phase with training 
data and application phase with new images [6]. This method is still being 
studied but is more complex than the neural network [2].

2.2 Image stylization using convolutional neural networks

An extension to the example-based method would be the texture synthesis using 
convolutional neural networks developed by Leon Gatys [14]. This is the 
pre-trained CNN used in the paper, which is also known as VGG-19. Gram matrices 
are introduced on the feature maps to transfer the representations into a 
stationary feature space [14]. There are also other works that have used 
convolutional neural networks for image stylization [10, 15]. In the process of 
image stylization, there is a trade-off between content and style matching, 
which is the content loss [10]. Furthermore, there are some limitations of 
using VGG networks due to the zero-padding used for convolutions in VGG network 
[14]. Overall, convolutional neural networks have a better performance than 
previous works on image stylization [21].

References for intro and related work:
[1] Haeberli, P. (1990). Paint by numbers: abstract image representations. SIGGRAPH Comput Graph, 24 (4), pp. 207-214. doi: 10.1145/97880.97902
[2] Hertzmann, A. (2018). Image Stylization: History and Future (Part1). Adobe Research. Viewed at: https://research.adobe.com/news/image-stylization-history-and-future/
[3] Hertzmann, A. (1998). Painterly rendering with curved brush strokes of multiple sizes. Proceedings of the 25th annual conference on Computer graphics and interactive techniques, pp. 453-460. doi: 10.1145/280814.280951 
[4] Hertzmann, A. (2018). Image Stylization: History and Future (Part2). Adobe Research. Viewed at: https://research.adobe.com/news/image-stylization-history-and-future-part-2/
[5] Hertzmann, A. (2018). Image Stylization: History and Future (Part3). Adobe Research. Viewed at: https://research.adobe.com/news/image-stylization-history-and-future-part-3/
[6] Hertzmann, A., Jacobs, C. E., Oliver, N., Curless, B., Salesin, D. H. (2001). Image analogies. Proceedings of the 28th annual conference on Computer graphics and interactive techniques, pp. 327-340. doi: 10.1145/383259.383295 
[7] Hegde, S., Gatzidis, C., Tian, F. (2012). Painterly rendering techniques: a state‐of‐the‐art review of current approaches. Wiley Online Library. doi: 10.1002/cav.1435
[8] Kyprianidis, J.E., Collomosse, J., Wang, T., Isenberg, T. (2013). State of the art: a taxonomy of artistic stylization techniques for images and video. IEEE Trans Vis Comput Graphics, 19 (5), pp. 866-885. doi: 10.1109/TVCG.2012.160
[9] Karpathy, A. (2019). Convolutional Neural Networks (CNNs/ConvNets). Stanford, CS231n Convolutional Neural Networks for Visual Recognition. Viewed at:  http://cs231n.github.io/convolutional-networks/
[10] Krizhevsky, A., Sutskever, I., Hinton, G. E. (2012). ImageNet Classification with Deep Convolutional Neural Networks. NIPS proceedings. Viewed at: http://papers.nips.cc/paper/4824-imagenet-classification-with-deep-convolutional-neural-networ
[11] Liu, Y., Qin, Z., Wan, T., Luo, Z. (2018). Auto-painter: Cartoon image generation from sketch by using conditional Wasserstein generative adversarial networks. Neurocomputing, Volume 311, pp. 78-87. ISSN 0925-2312. doi: 10.1016/j.neucom.2018.05.045. 
[12] Liang, M., Hu, X. (2015). Recurrent Convolutional Neural Network for Object Recognition. IEEE Conference on Computer Vision and Pattern Recognition (CVPR), pp. 3367-3375. Viewed at: http://openaccess.thecvf.com/content_cvpr_2015/html/Liang_Recurrent_Convolutional_Neural_2015_CVPR_paper.html
[13] Litwinowicz, P. (1997). Processing images and video for an impressionist effect. Apple Computer, Inc., CA, DOI: 10.1145/258734.258893
[14] Gatys, L. A., Ecker, A. S., Bethge, M. (2015), Texture Synthesis using Convolutional Neural Networks. Advances in Neural Information Processing Sytems 28 (NIPS 2015), Viewed at: http://papers.nips.cc/paper/5633-texture-synthesis-using-convolutional-neural-networks
[15] Gatys, L. A., Ecker, A. S., Bethge, M. (2016). Image Style Transfer Using Convolutional Neural Networks. IEEE Conference on Computer Vision and Pattern Recognition (CVPR), pp. 2414-2423. Viewed at: https://www.cv-foundation.org/openaccess/content_cvpr_2016/html/Gatys_Image_Style_Transfer_CVPR_2016_paper.html
[16] Rosin, P., Collomosse, J. (2013). Image and Video-Based Artistic Stylisation. Springer. doi: 10.1007/978-1-4471-4519-6 
[17] Schmidhuber, J. (2014). Deep learning in neural networks: An overview. Neural Networks, Volume 61, pp. 85-117. ISSN 0893-6080. doi: 10.1016/j.neunet.2014.09.003.
[18] Semmo, A., Limberger, D., Kyprianidis, J. E., Döllner, J. (2016). Image stylization by interactive oil paint filtering. Computers & Graphics, Volume 55, pp. 157-171, ISSN 0097-8493, doi: 10.1016/j.cag.2015.12.001.
[19] Super Data Science (2018). Convolutional Neural Networks (CNN): Step 1 – Convolution Operation. Super Data Science. Viewed at: https://www.superdatascience.com/blogs/convolutional-neural-networks-cnn-step-1-convolution-operation/ 
[20] Strassmann, S. (1986). Hairy brushes. Proceedings of the 13th Annual Conference on Computer Graphics and Interactive Techniques 1986, 20(4), pp.225–232.
[21] Zhu, F., Yu, Y. (2018). Automatic Image Stylization using Deep Fully Convolutional Networks. The University of Hong Kong. Viewed at: https://arxiv.org/pdf/1811.10872.pdf

\section{Method}

\subsection{Deep Neural Network}

For the purpose of implementing and even extending neural style transfer algorithms,
it is not necessary to understand how a neural network is trained, because
we can use pretrained model for example vgg19 \cite{vgg19}.

It is necessary to point out that a deep neural network that neural style transfer
is concerend about
is just a sequence of functions
$F_1,F_2,F_3,...$
whose inputs and outputs are all multi-dimensional arrays, as illustrated in Figure \ref{f1f2f3}.
\begin{figure}
\center
\includegraphics[width=0.6\textwidth]{f1f2f3.pdf}
\caption{A deep neural network \label{f1f2f3}}
\end{figure}
Because the arrays have usually more than two numbers of dimensions
they are not really matrices.
The arrays are called tensors (as in TensorFlow or torch.Tensor in PyTorch) 
but actual tensor algebra in a mathematical sense is rarely relevant.
The functions need to be differentiable, which is handled automatically by modern 
deep learning frameworks.
Other than that we can treat the functions as blackboxes because we are 
not concerned of training the model.
A ``layer'' in a deep neural network refers to a few consecutive functions together.
It is not particularly necessary to consider ``layers'' in this work. 
Though it is important to note that in our illustrations there are functions instead of layers.

\subsection{vgg19 model}
For the vgg19 model that most neural style transfer implementations available are based on,
The inputs and outpus of each function are all arrays with 4 dimensions.
The word ``dimension'' here may mean something slightly different from what ``dimension'' means in
for example ``$3$-dimension vector''. 
Some people call the number of dimensions ``rank'' which would then raise another confusion with the rank of matrices.
We give an illustration for the vgg19 model in Figure \ref{vgg512}
\begin{figure}
\center
\includegraphics[width=\textwidth]{vgg512.pdf}
\caption{The VGG19 model with an $512\times512$ RGB image as its input \label{vgg512}}
\end{figure}

The 1st dimension is the ``batch'' dimension, which means that you can put several images through the
series of functions.
In neural style transfer implementations it is usually just one image each time.
So the 1st dimension is always 1, even between the $F$ functions, throughout this work.

The 2nd dimension is the ``feature'' dimension. For a raw RGB image, it is 3.
Intermediate arrays between, for example, $F_3$ and $F_4$, usually 
have a size much larger than 3, such as 64 or 128.

The 3nd and 4th dimensions are just spatial locations. 
For a $512\times512$ image they are 512 and 512. 
For intermediate arrays, these 2 dimensions are sometimes scaled down
to half or even $\frac{1}{4}$ of the sizes of the image.

The functions $F_1,F_2,F_3,...$ now have meaningful names such as relu\_1, conv\_2 now.
[[[[Not actually necessary to know but insert your elaborations here]]]]

The arrays after the initial image are called \emph{feature maps}, because their value means
``how much a feature x exists at position y and x''.
Take a $1\times64\times512\times512$ feature map for example. There are 64 types of features.
Its [1, 17, 111, 222] element, would then refer to the extent that the 17th type feature exists
at the position of the 222th column of the 111th row. This ``extent'' could also be negative.

\subsection{Vanilla neural style transfer}

The vanillay neural style transfer algorithm \cite{nst} is 
structured as iteratively solving an optimisation problem.
The argument to optimise is the image as a multidimensional array,
of size $1\times3\times512\times512$ for example.
The interative solver can be many, but in our case it is 
L-BFGS [[[[Maybe elaborate on L-BFGS here]]]]. 
The initial value of the argument could be either white noise or the content image,
but the latter appeared to make the optimisation much easier.

The key issue here is still how to structure the optimisation target.
The value to minimise is a linear combination of a ``style loss'' 
and ``content loss''.
Although these appeared to be two weights,
actually only their ratio mattered.
We simply fixed the weight of the content loss to 1 and
leave the weight of the style loss as a adjustable hyperparameter $k$.
\begin{align}
\argmin\limits_{\text{image}} (kL_\text{style}+L_\text{content})
\end{align}

The content loss $L_\text{content}$ is the easier one to explain of the two losses.
It is the mean square error between the outputs of a certain function in the
vgg19 taking the current argument image and the content image as inputs respectively.
We say ``a certain'' function because the output of which function to choose is
adjustable.
A typical choice is the output of \verb|conv_4| function. See Figure \ref{contentloss}.
\begin{figure}
\center
\includegraphics[width=\textwidth]{contentloss.pdf}
\caption{The content loss \label{contentloss}}
\end{figure}

The style loss $L_\text{content}$ can actually be a sum of several style losses, or sometimes just one.
For each style loss, 
both the current image under optimisation and the style image are passed through the vgg19 model.
The two outputs of a certain function in the model are extracted, 
and converted to Gram matrices. We will elabrate on what Gram matrices are later.
For now a Gram matrix is just a square matrix computed from a multidimensional array (i.e. a feature map).
For example, the Gram matrix (in the style transfer algorithm) of a $1\times64\times512\times512$ array
is a $64\times64$ matrix.
The point is that a Gram matrix describes the distribution of features
in the image without the spatial information of where the features are.
Gram matrices describe the distribution with correlation between features.
The several style losses are computed with the outputs of different functions.
Otherwise, each of the several style losses are computed in exactly the same way.
Again, the outputs of which functions in the model to derive style losses from is really adjustable.
A nice default we used was \verb|conv_1|,\verb|conv_2|,...,\verb|conv_5|. We illustrate
the style losses derived from the output of \verb|conv_1| in Figure $\ref{styleloss}$.

\begin{figure}
\center
\includegraphics[width=\textwidth]{styleloss.pdf}
\caption{The style loss derived from the output of conv\_1 \label{styleloss}}
\end{figure}

\subsection{Gram matrix}
[[[[Elaborate on gram matrix here]]]]

\subsection{Removing the spatial information in other ways}
We think that the most brilliant aspect of the neural style transfer algorithm,
apart from the general structure, is the use of Gram matrices to compare 
distributions of features in two images, so that where the features are does not matter.
It gives really nice insight on what ``artistic style'' means.

Now we wonder, could we achieve similar results with other ways to compare 
distribution of features?
One approach would be to consider the \emph{comparison} as a whole to find 
substitutes. 
For example, minimising the mean square errors between two Gram matrices could be seen as 
a special case for minimising a Maximum Mean Discrepancy (MMD), which means other Maximum Mean Discrepacies
could be used \cite{MMD}.

[[[[Maybe Describe MMD here]]]]

For us, it seemed more natural to consider substitutes for the process itself of 
transforming a feature map into a Gram matrix. 
That is, we will consider other
mappings that remove the spatial information or where features are in a feature map but retain
the distribution of features, like a histogram.
Such functions need to be differentiable as well, otherwise they would render the 
optimisation infeasible.

We will call such functions ``\emph{dislocators}'' in the remaing part of this work.
Unfortunately, the word ``dislocate'' would give the connotation of moving something to somewhere else.
Nevertheless, the word ``locator'' means something that provides spatial information,
so calling somthing that removes spatial information a ``dislocator'' is still reasonable.

Other ``dislocator'' functions whose output somehow describe the 
distribution of features in their input should then work as substitutes
for computing Gram matrices.
The first few simplistic ``dislocator'' functions we tried are listed here.
\begin{enumerate}
\item Scaled sum of the responses all over the place for each feature.
That is, for an $1\times64\times256\times256$ feature map,
we compute an $1\times64$ array where each element is the sum of the
$256\times256$ elements from the input.
The whole array is then scaled to one over the total number of elements of the input feature map.
\item Scaled sum of element-wise squares of the responses over the place for each feature.
Other than an additional element-wise squaring, it is identical to the previous one.
\item Scaled sum of absolute values of the responses over the place for each feature.
\end{enumerate}

To find more dislocators, let us consider \emph{how to describe a distribution} in general?
The obvious answer is then sample statistics or estimators, like (sample) mean, (sample) standard deviation, 
(sample) skewness, (sample) kurtosis and any moments.
Unfortunately, we found that higher moments too slow to compute for our experiments,
so we ended up with the following (combinations) of sample statistics.
\begin{enumerate}
\item Mean and standard deviation of each type of feature. 
That is, given a $1\times64\times256\times256$ feature map, 
we compute an $1\times64\times2$ array.
Standard deviation is favored over variance because standard deviations 
are supposedly of the same unit of measure as the mean, 
while variances are of a squared unit of measure.
\item Mean, standard deviation and 4th central moment to the power of $1/4$. 
That is, a $1\times64\times3$ array for a $1\times64\times256\times256$ feature map.
\item Mean, standard deviation, 4th central moment to the power of $1/4$, 
and 6th central moment to the power of $1/6$.
\end{enumerate}
We have found by experiment that odd-order moments did not work.
6th central moments were already much slower to compute in practice 
than Gram matrices.
So we could only stop at the 6th moment due to limited experimental capacity.

Unfortunately, straight forward histograms were not computationally feasible.
Although the autograd facility could handle histograms in theory, we were not
able to make histograms work in practice. Similarly, sorting or order statistics
like medians were found to be computationally infeasible in practice.

\section{Experiments and Discussion}

\subsection{Reproducing the original result}
[[[[Fill the experiment part in demo here to show that we have reproduced result.
Cite the original tutorial]]]] \cite{tut}

\subsection{Varying the extraction points for losses}

We will now show that within the general optimisation framework,
steps other than the Gram matrix, 
such as the outputs of which feature maps to derive the style 
losses from are not very essential.

\subsection{Varying the dislocator function}

Now let us see something more interesting.

\section{Conclusion and Future Work}

\begin{thebibliography}{8}
\bibitem{nst}
Gatys, L., Ecker, A., \& Bethge, M. (2016). A neural algorithm of artistic style. Journal of Vision, 16(12), 326. doi:10.1167/16.12.326
\bibitem{MMD}
Li, Y., Wang, N., Liu, J., \& Hou, X. (2017). Demystifying Neural Style Transfer. IJCAI.
\bibitem{interp}
Doshi-Velez, F., \& Kim, B. (2017). Towards A Rigorous Science of Interpretable Machine Learning.
\bibitem{pytorch}
Paszke, A., Gross, S., Chintala, S., Chanan, G., Yang, E., DeVito, Z., Lin, Z., Desmaison, A., Antiga, L., \& Lerer, A. (2017). Automatic differentiation in PyTorch.
\bibitem{tut}
Jacq, A. (n.d.). Neural Transfer Using PyTorch (W. Herring, Ed.). Retrieved from https://pytorch.org/tutorials/advanced/neural\_style\_tutorial.html
\end{thebibliography}
\end{document}